---
title: "Case Study 2 - Mental Health in the Tech Industry"
output:
  html_document: default
---
### Loading Data and General Manipulation
```{r}
library(tidyverse)
library(readr)
library(skimr)
library(lubridate)
library(SimDesign)
library(janitor)
library(RSQLite)
library(knitr)
library(scales)

con <- dbConnect(drv=RSQLite::SQLite(), dbname="mental_health.sqlite")
df <- dbSendQuery(con, "SELECT * FROM MASTER_DATA")
MASTER_DATA <- dbFetch(df)

MASTER_DATA <- rename(MASTER_DATA,
                      "ANSWER" = "AnswerText",
                      "YEAR" = "SurveyID",
                      "USER_ID" = "UserID",
                      "QUESTION_ID" = "Questionid",
                      "QUESTION" = "Questiontext")
skim_without_charts(MASTER_DATA)
```
### Gathering General Information about the survey
```{r}
glimpse(unique(MASTER_DATA$QUESTION ))
```

### Getting the IDs of the Questions for selection
```{r}
MASTER_DATA%>%
  group_by(MASTER_DATA$QUESTION_ID,QUESTION)%>%
  summarise()
```


### Selection of the core questions to be used for analysis
```{r}
selected_question_IDs <- c(1, 2, 3, 5, 117, 33, 89, 96, 48, 49, 65)
MASTER_DATA <- MASTER_DATA%>%
  filter(QUESTION_ID %in% selected_question_IDs)
unique(MASTER_DATA$QUESTION_ID)
#Count of answers per selected question
table(MASTER_DATA$QUESTION)
```


### User Age Statistics
```{r}
AGE_DATA <-
  filter(MASTER_DATA,MASTER_DATA$QUESTION_ID == 1)
AGE_DATA$ANSWER <- as.numeric(AGE_DATA$ANSWER)
for (q in AGE_DATA$QUESTION_ID) {
  if(q == 1){
    AGE_DATA$ANSWER[AGE_DATA$ANSWER>80 | AGE_DATA$ANSWER < 18] <- NA
    }
  }

min_age <- setNames(aggregate(AGE_DATA$ANSWER~AGE_DATA$QUESTION,FUN = "min"),c("Question","Min_Value"))
max_age <- setNames(aggregate(AGE_DATA$ANSWER~AGE_DATA$QUESTION,FUN = "max"),c("Question","Max_Value"))
average_age <- setNames(aggregate(AGE_DATA$ANSWER~AGE_DATA$QUESTION,FUN = "mean"),c("Question","Mean"))
min_age
max_age
average_age
```

### Gender of the Survey users
```{r}
#Dividing the gender to Male/Female/Other

GENDER_DATA <- MASTER_DATA%>%
  filter(QUESTION_ID == 2 )

#Summing down the gender categories into 3
GENDER_DATA$ANSWER[GENDER_DATA$ANSWER != "Male" & GENDER_DATA$ANSWER != "Female"] <- "Other"

#breakdown of Genders
table(GENDER_DATA$YEAR,GENDER_DATA$ANSWER)

#Plotting the gender over the years
GENDER_DATA %>%
  group_by(ANSWER,YEAR)%>%
  summarise(number = n())%>%
  ggplot(aes(x = YEAR, y = number , fill = ANSWER, group = ANSWER, color = ANSWER)) + geom_line(linetype = 1, color = "red") + geom_point() + labs(y = "Count of Gender", title = "Count of type of Gender per Year")

```

### Countries where the users live in
```{r}
COUNTRY_DATA <- MASTER_DATA %>% 
  filter(QUESTION_ID == 3)

#Removed the Countries marked as "-1" and "Other"
COUNTRY_DATA <- subset(COUNTRY_DATA,ANSWER != -1 & ANSWER != "Other")

#replacing "United States" by "United States of America" and "Bahamas, The" with "The Bahamas"
COUNTRY_DATA$ANSWER[COUNTRY_DATA$ANSWER == "United States"] <- "United States of America"
COUNTRY_DATA$ANSWER[COUNTRY_DATA$ANSWER == "Bahamas, The"] <- "The Bahamas"

table(COUNTRY_DATA$ANSWER,COUNTRY_DATA$YEAR)

COUNTRY_DATA%>%
  group_by(ANSWER,YEAR)%>%
  summarize(number = n())%>%
  arrange(desc(number),YEAR,ANSWER)
```


### How much of the users are self employed
```{r}
#Taking out the answers that are -1 as this is a faulty data
EMP_DATA <- MASTER_DATA%>%
  filter(QUESTION_ID == 5)

EMP_DATA <- subset(EMP_DATA, ANSWER != -1)

for (a in EMP_DATA$ANSWER) {
  if(a == "1" | a == 1){
    EMP_DATA$ANSWER[EMP_DATA$ANSWER == 1 | EMP_DATA$ANSWER == "1"] <- "Yes"
  }
  else{ 
    EMP_DATA$ANSWER[EMP_DATA$ANSWER == 0 | EMP_DATA$ANSWER == "0"] <- "No"}
}
table(EMP_DATA$YEAR,EMP_DATA$ANSWER)

#Visualizing the count of the Answers per year
EMP_DATA%>%
  group_by(ANSWER,YEAR)%>%
  summarize(Count = n())%>%
  
ggplot(aes(x = YEAR, y = Count, fill = ANSWER, color = ANSWER)) + geom_col(position = "dodge") + geom_point() + geom_line(linetype = 1, color = "red")


#Visualizing the percentage per answer per year
EMP_DATA2 <- EMP_DATA%>%
  mutate(PercTot = round(1 / length(ANSWER),digits = 5))

p2 <- ggplot(data = EMP_DATA2,aes(x = YEAR, y = PercTot, fill = ANSWER, color = ANSWER))
p2 + 
  geom_bar(stat = "identity", #confusing subject, when plotting pivoted data, always leave "identity"
            position = "fill",
            width = 0.6,      # how wide (0-1) the bars are
            color = "black",  # the outline color of the bars
            size = 0.5,         # the thickness of the outline
            alpha = 0.9) +    # the opaqueness of the fill colors
   theme_minimal()+           # the background
   theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 0.8)) +                         # legend position
   labs(x= "Year", y = "Percent", title = "Self-Employment % per Year") +  # axis labels and footnote
   scale_x_discrete(limits = EMP_DATA2$YEAR) + # filter only certain factors
   geom_text(aes(label = PercTot), position = position_fill(vjust = 0.5), size = 3)  # the labels

```

### What are the job positions of the users
```{r}
POS_DATA <- MASTER_DATA%>%
  filter(QUESTION_ID == 117)
table(POS_DATA$ANSWER,POS_DATA$YEAR)


POS_DATA%>%
  group_by(ANSWER,YEAR)%>%
  summarize(Count = n())%>%
  ggplot(aes(x = "", y = Count, fill = ANSWER)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + labs(title = "Count of Job positions")
```


### How many people currently have mental disorders?
```{r}
MDO_DATA <- MASTER_DATA%>%
  filter(QUESTION_ID == 33)


#Replace "Possibly" with "Maybe"
for (a in MDO_DATA$ANSWER) {
  if(a == "Possibly"){
    MDO_DATA$ANSWER[MDO_DATA$ANSWER == "Possibly"] <- "Maybe"
  }
}
table(MDO_DATA$YEAR,MDO_DATA$ANSWER)

#Visualizing the Count of the answers
MDO_DATA%>%
  group_by(ANSWER,YEAR)%>%
  summarize(Count = n())%>%
  ggplot(aes(x = YEAR, y = Count, fill = ANSWER, color = ANSWER)) + geom_col(position = "dodge") + geom_point() + geom_line(linetype = 1, color = "red")
  

# Visualizing the percentage per answer per year
MDO_DATA2 <- MDO_DATA%>%
  mutate(PercTot = round(1 / length(ANSWER),5))

ggplot(data = MDO_DATA2,aes(x = MDO_DATA2$YEAR, y = MDO_DATA2$PercTot, fill = ANSWER, color = ANSWER)) +
  geom_bar(stat = "identity",
            position = "fill",
            width = 0.5,  
            color = "black",
            size = 0.5,
            alpha = 0.9) +
   theme_minimal() +
   theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 0.8)) +
   labs(x= "Year", y = "Percent", title = "% of Mental Disorders per Year") +  
   scale_x_discrete(limits = MDO_DATA2$YEAR) +
   geom_text(aes(label = PercTot), position = position_fill(vjust = 0.5), size = 3) +
   labs(title = "Mental disorder count per year")


```


### The race of the users
```{r}
RACE_DATA <- MASTER_DATA %>%
  filter(QUESTION_ID == 89)

for (a in RACE_DATA$ANSWER) {
  if(a == -1){
    RACE_DATA$ANSWER[RACE_DATA$ANSWER == -1] <- "Prefer not to say"
  }
}

RACE_DATA <- RACE_DATA%>%
  filter(ANSWER != -1)

table(RACE_DATA$ANSWER,RACE_DATA$YEAR)

RACE_DATA%>%
  group_by(ANSWER,YEAR)%>%
  summarize(Count = n())%>%
  ggplot(aes(x = "", y = Count, fill = ANSWER)) + geom_bar(width = 6, stat = "identity") + coord_polar("y", start=0) + labs(title = "Race of the Participants")

```


### Does your employer provide resources to learn more about mental health issues and how to seek help?
```{r}
HELP_DATA <- MASTER_DATA%>%
  filter(QUESTION_ID == 96)
table(HELP_DATA$ANSWER,HELP_DATA$YEAR)

HELP_DATA%>%
  group_by(ANSWER,YEAR)%>%
  summarize(Count = n())%>%
  ggplot(aes(x = "", y = Count, fill = ANSWER)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + labs(title = "Does your employer provide resources regarding mental health issues?")
```


### If you have a mental health disorder, how often do you feel that it interferes with your work when being treated effectively?
```{r}
TREATED_DATA <- MASTER_DATA%>%
  filter(QUESTION_ID == 48)

table(TREATED_DATA$ANSWER,TREATED_DATA$YEAR)

#Count of answers per user per year
TREATED_DATA%>%
  group_by(ANSWER,YEAR)%>%
  summarize(Count = n())%>%
  ggplot(aes(x = YEAR, y = Count, fill = ANSWER, color = ANSWER)) + geom_col(position = "dodge") + geom_point() + geom_line(linetype = 1, color = "black")


#Percentage of each answer per year
TREATED_DATA2 <- TREATED_DATA%>%
  mutate(PercTot = round(1 / length(ANSWER),5))

ggplot(data = TREATED_DATA2,aes(x = TREATED_DATA2$YEAR, y = TREATED_DATA2$PercTot, fill = ANSWER, color = ANSWER)) +
  geom_bar(stat = "identity",
            position = "fill",
            width = 0.5,  
            color = "black",
            size = 0.5,
            alpha = 0.9) +
   theme_minimal() +
   theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 0.8)) +
   labs(x= "Year", y = "Percent", title = "% of Interference of the mental Disorder in work when treated") +  
   scale_x_discrete(limits = TREATED_DATA2$YEAR) +
   geom_text(aes(label = PercTot), position = position_fill(vjust = 0.5), size = 3)

```


### If you have a mental health disorder, how often do you feel that it interferes with your work when not being treated effectively (i.e., when you are experiencing symptoms)?
```{r}
nTREATED_DATA <- MASTER_DATA%>%
  filter(QUESTION_ID == 49)
table(nTREATED_DATA$ANSWER, nTREATED_DATA$YEAR)

#Count of the answers per year
nTREATED_DATA%>%
  group_by(ANSWER,YEAR)%>%
  summarize(Count = n())%>%
  ggplot(aes(x = YEAR, y = Count, fill = ANSWER, color = ANSWER)) + geom_col(position = "dodge") + geom_point() + geom_line(linetype = 1, color = "black")


#Percentage of each answer per year
nTREATED_DATA2 <- nTREATED_DATA%>%
  mutate(PercTot = round(1 / length(ANSWER),5))

ggplot(data = nTREATED_DATA2,aes(x = nTREATED_DATA2$YEAR, y = nTREATED_DATA2$PercTot, fill = ANSWER, color = ANSWER)) +
  geom_bar(stat = "identity",
            position = "fill",
            width = 0.5,  
            color = "black",
            size = 0.5,
            alpha = 0.9) +
   theme_minimal() +
   theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 0.8)) +
   labs(x= "Year", y = "Percent", title = "% of Interference of the mental Disorder in work when not treated") +  
   scale_x_discrete(limits = nTREATED_DATA2$YEAR) +
   geom_text(aes(label = PercTot), position = position_fill(vjust = 0.5), size = 3)

```


### Overall, how much importance does your employer place on mental health?
```{r}
IMP_DATA <- MASTER_DATA%>%
  filter(QUESTION_ID == 65)

#None = -1 till 3
#Some = 4 till 7
#Very much = 8 till 10

none = c(-1,0,1,2,3)
some = c(4,5,6,7)
Very = c(8,9,10)

for (a in IMP_DATA$ANSWER) {
  IMP_DATA$ANSWER[IMP_DATA$ANSWER %in% none] <- "NONE"
  IMP_DATA$ANSWER[IMP_DATA$ANSWER %in% some] <- "SOME"
  IMP_DATA$ANSWER[IMP_DATA$ANSWER %in% Very] <- "Very Much"
}
table(IMP_DATA$ANSWER,IMP_DATA$YEAR)


#Importance count per user's employers per year
IMP_DATA%>%
  group_by(ANSWER,YEAR)%>%
  summarize(Count = n())%>%
  ggplot(aes(x = YEAR, y = Count, fill = ANSWER, color = ANSWER)) + geom_col(position = "dodge") + geom_point() + geom_line(linetype = 1, color = "black")


#Percentage of importance per users' employers per year
IMP_DATA2 <- IMP_DATA%>%
  mutate(PercTot = round(1 / length(ANSWER),5))

ggplot(data = IMP_DATA2,aes(x = IMP_DATA2$YEAR, y = IMP_DATA2$PercTot, fill = ANSWER, color = ANSWER)) +
  geom_bar(stat = "identity",
            position = "fill",
            width = 0.4,  
            color = "black",
            size = 0.5,
            alpha = 0.9) +
   theme_minimal() +
   theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 0.8)) +
   labs(x= "Year", y = "Percent", title = "% of Importance given By Employers on Mental Health") +  
   scale_x_discrete(limits = IMP_DATA2$YEAR) +
   geom_text(aes(label = PercTot), position = position_fill(vjust = 0.5), size = 3)

```


